On the Physics of Radio Halos in Galaxy Clusters: Scaling Relations and Luminosity Functions

### 2013-09-16 paper2.tex
New Fabio Version

  from the MultiDark $N$-body $\Lambda$CDM simulation.} 

@@

  a physical solution to this apparent contradiction. {\bf Assuming a plausible fraction
  of radio-loud clusters, fixed at all redshifts, we demonstrate the unique 

@@

  LOFAR Tier~1 survey) to detect $\sim600\times\rmn{\emph{loud~fraction}[\%]}/10$ 
  hadronically-generated (giant and mini) radio halos and to probe the underlying 
  physics of radio halos. This shows the potentially of future radio surveys in determining 

@@

\cite{2011A&A...527A..99E} tried to asses these problems of the classical hadronic 
model by analyzing CR transport processes within a cluster. While CR advection with 
turbulently driven flows results in centrally enhanced CR profiles, the propagation in 
form of CR streaming and diffusion produces a flattening of CR profiles. Hence, 
different CR transport phenomena may also account for the observed bimodality 
of the radio luminosity in the hadronic model and may also explain the spectral 
features observed in some clusters. {\bf This has been recently confirmed 
by \cite{2013arXiv1303.4746W} that developed a more rigorous treatment. 
In particular, they show that when considering turbulent
damping, CRs can stream at super-Alfvenic velocities, which is crucial
for CR transport to be of relevant magnitude.} 
Note that these phenomena were not considered in earlier analytical works 
(e.g., \citealp{2004A&A...413...17P}) as well as in previous hydrodynamic 
simulations (e.g., \citealp{2001ApJ...562..233M, 
2008MNRAS.385.1211P, 2010MNRAS.409..449P}).
{\bf A satisfactory theory of CR transport in clusters does not exist yet, 
however CR streaming and diffusion may represent an intriguing solution
for the issues of the classical hadronic model.}

@@

An observational test that is able to disentangle between the hadronic and
re-acceleration models is the gamma-ray emission resulting from neutral pion
decays, a secondary product of the hadronic CR interaction with protons of the
ICM, which is not predicted by the re-acceleration model. Such observational
efforts have been undertaken in the last few years (for space-based cluster
observations in the GeV-band, see \citealt{2003ApJ...588..155R,
  2010JCAP...05..025A,2010ApJ...717L..71A,2012AAS...21920701Z,2012arXiv1207.6749H}; for ground-based
observations in the energy band $\gtrsim100$~GeV, see
\citealt{2006ApJ...644..148P, 2008AIPC.1085..569P,
  2009arXiv0907.0727T,2009A&A...495...27A, 2009arXiv0907.3001D,
  2009arXiv0907.5000G, cangaroo_clusters, 2009ApJ...706L.275A,
  2010ApJ...710..634A, 2011arXiv1111.5544M,2012...VERITAS,2012A&A...545A.103H})
without being able to detect cluster gamma-ray emission. Current gamma-ray 
limits enable us to constrain the average CR-to-thermal pressure to be less than a few
percent, and the maximum CR acceleration efficiency at structure formation shocks 
to be $<50\%$. Alternatively, this may indeed suggest the presence of non-negligible 
CR transport processes into the outer cluster regions. 
%significant constraints on the relative CR pressure contribution. 
%The long observational campaign of the Perseus galaxy cluster
%performed by the Major Atmospheric Gamma-ray Cherenkov (MAGIC) telescopes
%constrains the cluster average CR-to-thermal pressure to be less than a few
%percent \citep{2010ApJ...710..634A,2011arXiv1111.5544M}. Comparing the
%corresponding upper limits to predictions of simulations by
%\cite{2010MNRAS.409..449P} constrains the maximum CR acceleration efficiency at
%structure formation shocks to be $<50\%$.  Alternatively, this may indeed
%suggest the presence of non-negligible CR transport processes into the outer
%cluster regions as suggested by \cite{2011A&A...527A..99E}.  Constraints at the
%same level are confirmed by the \emph{Fermi}-Large Area Telescope (LAT) data on
%the Coma cluster \citep{2011arXiv1105.3240P,2012AAS...21920701Z,
%2012...VERITAS,2012arXiv1207.6749H}. To summarize, gamma-ray observations are 
%not yet sensitive enough to severely challenge hadronic models of RHs.

@@

{\bf The main scope of this work is to improve the classical hadronic model with
the inclusion of possible CR transport phenomena, trying in this way to solve
its main issues when comparing to observations and to make predictions for
future radio surveys. The outline is as follows. In Section~\ref{sec:2.3}, we 
construct a model for the CR proton distribution in clusters that merges results 
of hydrodynamic cluster simulations and an analytical model for microscopic 
CR transport processes. In Section~\ref{sec:3}, we model observed surface 
brightness profiles of individual RHs within the hadronic scenario and explore the 
allowable parameter space for CRs and magnetic fields. Motivated by immanent 
challenges to explain the extent of giant radio halos within the hadronic model, 
we suggest a new hybrid leptonic/hadronic scenario meant to unify the apparently 
distinct classes of giant radio halos, mini halos, and steep spectrum radio sources in 
Section~\ref{sec:discussion_hadronic}. In Section~\ref{sec:4}, we apply our
extended hadronic model to a cosmologically complete mock galaxy cluster catalog 
built from the MultiDark $N$-body $\Lambda$CDM simulation
in Zandanel, Pfrommer \& Prada (2013; hereafter Paper~I). 
We compare the resulting modeled radio-to-X-ray and radio-to-SZ
scaling relations to current observations and show how they vary for different
choices of our CR and magnetic parametrizations. In Section~\ref{sec:5}, we show
the radio luminosity functions, compare them to current observational
constraints and provide predictions for the hadronically-generated
RHs for the LOFAR cluster survey. Finally, in Section~\ref{sec:6}, we present 
our conclusions.} In this work, the cluster mass
$M_{\Delta}$ and radius $R_{\Delta}$ are defined with respect to a density that
is $\Delta=200$ or $\Delta=500$ times the \emph{critical} density of the
Universe. We adopt density parameters of $\Omega_{\rmn{m}}=0.3$,
$\Omega_{\Lambda}=0.7$ and today's Hubble constant of $H_0 = 100 \times
h_{70}$~km~s$^{-1}$~Mpc$^{-1}$ where $h_{70} = 0.7$.

@@

{\bf To construct such a model, we have to generalize the approach proposed by
\citet{2011A&A...527A..99E}, which uses a $\beta$-profile gas parameterizatio, 
in order to account for different kind of ICM gas profiles, such as our generalized 
Navarro-Frank-White (GNFW) profiles derived in Paper~I from X-ray observations.} 
We also have to include the cluster mass-scaling of the
CR normalization obtained from simulations \citep{2010MNRAS.409..449P}. While
details are given in Appendix~\ref{app:B}, we show below the main steps. When
turbulent advection completely dominates the CR transport, the CR normalization
can be written as \citep{2011A&A...527A..99E}
\begin{equation}
C_{\rmn{adv}}(R)=C_{0} \left( \frac{P_{\rmn{th}}(R)}{P_{\rmn{th},0}} \right)^{\frac{\beta_{\rmn{CR}}}{\gamma}} = 
C_{0} \eta(R)^{\beta_{\rmn{CR}}},
\label{eq:Csimple_1}
\end{equation} 
where $P_{\rmn{th}}$ is the thermal pressure, $\beta_{\rmn{CR}}=(\alpha+2)/3$, $\gamma=5/3$, 
and we introduced the advective CR profile $\eta(R)=(P_{\rmn{th}}(R)/P_{\rmn{th},0})^{1/\gamma}$. Solving the continuity
equation for CRs, \citet{2011A&A...527A..99E} derive the CR density profile,
\begin{equation}
\rho_{\rmn{CR}}(R) = \rho_{\rmn{CR},0} \eta(R) \rmn{exp} \left( \frac{R}{R_{*}} \right) \, ,
\label{eg:rhoCR_1}
\end{equation} 
where $R_{*}=\gamma_{\rmn{\rmn{tu}}}R_{\rmn{c}}$ and $R_{\rmn{c}}$ is the characteristic 
radius, of order the core radius, at which the turbulence is supposed to be injected.
Now, we introduce the \emph{semi-analytical} mass-dependent normalization of the
CR profile of \cite{2010MNRAS.409..449P} such that
\begin{equation}
\eta(R) = \left( \frac{C_{\rmn{adv}}(R)}{C_0} \right)^{1/\beta_{\rmn{CR}}} \to
\left( \frac{C_{\rmn{extended}}(R)}{C_0} \right)^{1/\beta_{\rmn{CR}}} \, ,
\label{eq:eta}
\end{equation} 
which effectively redefines $C_{\rmn{adv}}(R)$ by that of our extended model, i.e.,
\begin{equation}
C_{\rmn{extended}}(R) =  \tilde{C}(R)\, \frac{\rho_{\rmn{gas}}(R)}{m_\rmn{p}} \frac{T(R)}{T_0}.
\label{eq:Cf}
\end{equation} 
Here, $\rho_{\rmn{gas}}$ is the ICM gas density and $\tilde{C}(R)$ is the normalization 
CR profile of equation~(22) of \cite{2010MNRAS.409..449P}. We additionally account for the 
temperature decline toward the cluster periphery, $T(R)$, given by the fit to the universal
temperature profile obtained from cosmological hydrodynamical simulations
\citep{2007MNRAS.378..385P,2010MNRAS.409..449P} and deep {\em Chandra} X-ray
observations \citep{2005ApJ...628..655V}. Eventually, the CR profile in our
extended model is $C(R)=C_{0}(\rho_{\rmn{CR}}(R) /
\rho_{\rmn{CR},0})^{\beta_{\rmn{CR}}}$ within $R_{\pm}$ of
equation~(\ref{eq:Rpm}), with $\rho_{\rmn{CR}}$ defined by
equation~(\ref{eg:rhoCR_1}) where $C_{\rmn{extended}}$ enters through our
redefinition of $\eta$, and $C(R) = C(R_{\pm})$ for $R > R_{+}$ and $R < R_{-}$,
respectively.

@@

In the right panel of Fig.~\ref{fig:CRFinalModel}, we compare our extended model
profile with the semi-analytical advection-only case (adopting our GNFW gas
profiles and the outer temperature decrease to the model proposed by
\citealp{2010MNRAS.409..449P}) and with the exact analytical solution as in
\citet{2011A&A...527A..99E}, but for our GNFW profiles (Appendix~\ref{app:B}
for details).  The profiles are normalized at $0.1 R_{200}$. In the case of
dominant advective CR transport, our extended model compares nicely with the
semi-analytical model derived from cosmological cluster simulations
\citep{2010MNRAS.409..449P}.  The main differences between our extended model (and
the semi-analytical model) on the one side and the analytical solution on the
other side is the inclusion of the simulation-based ``reference'' profile
$\tilde{C}$ for the advection-only case and the universally observed temperature
drop towards the outskirts of clusters. Note that the profiles in our extended
model are generally more centrally peaked in comparison to the analytical GNFW
case, which is due to the enhanced radiative cooling in the
\citet{2010MNRAS.409..449P} simulations that did not account for AGN
feedback. Thanks to the flexible parametrization in our model, this can be
easily counteracted by changing $\gamma_{\rmn{tu}}$ and $ \alpha_{\rmn{B}}$ 
(representing the magnetic field radial decline, see next section), however, at
the expense that these parameters are now degenerate with our assumptions  on the
CR profile in the advection-dominated regime and other possible effects that we
are not considering, such as cluster asphericity.

@@

For the purpose of this section, we adopt the
measured gas and temperature profiles derived from X-ray observations of each cluster.  
Our extended model includes an overall normalization $g_{\rmn{CR}}$ of the CR
distribution function and the hadronically-induced non-thermal emission 
(Appendix~\ref{app:A}). Note that
this parameter can be interpreted as a functional that depends on the
\emph{maximum CR acceleration efficiency}, $g(\zeta_{\rmn{p,max}})$,
\citep{2010MNRAS.409..449P} but \emph{only} for
$\gamma_{\rmn{tu}}\gtrsim100$.

@@

The gamma-ray flux (Appendix~\ref{app:C}) within $R_{200}$ for the parameter
combination $\gamma_{\rmn{tu}} = 2$ and $\alpha_{\rmn{B}}=0.5$ and for energies
above 100~MeV (100~GeV) is $F_{\gamma} = 2.4 \times 10^{-9}$ ($8.7 \times
10^{-13}$) cm$^{-2}$~s$^{-1}$. This is below the the gamma-ray limit recently
set with the \emph{Fermi} data of $F_{\gamma, UL} (>100~\rmn{MeV}) \approx 3.2
\times 10^{-9}$~cm$^{-2}$~s$^{-1}$ \citep{2012...VERITAS}.  Note that for
slightly higher values of $\gamma_{\rmn{tu}}$, i.e., a more centrally
concentrated CR distribution, the radio and gamma-ray yield would be increased
(assuming CR number conservation). However, in order to match the observed radio
synchrotron profiles, we have to decrease the CR normalization (parametrized by
$g_{\rmn{CR}}$). This causes the associated gamma-ray flux also to be reduced to
a level that is low enough to easily circumvent the gamma-ray constraints. E.g.,
for the parameter combination $\gamma_{\rmn{tu}} = 3$ and
$\alpha_{\rmn{B}}=0.4$, we obtain $F_{\gamma} (>100~\rmn{MeV}) = 1.3 \times
10^{-9}$~cm$^{-2}$~s$^{-1}$ (and $F_{\gamma} (>100~\rmn{GeV}) = 4.9 \times
10^{-13}$~cm$^{-2}$~s$^{-1}$).  We caution that in principle, CR streaming
should cause the CR spectrum to steepen \citep{2013arXiv1303.4746W}.  
This may then considerably weaken these constraints as a result of the convex spectral 
curvature since the gamma-ray emission probes the high-energy tail of the CR distribution 
that is suppressed in this picture in comparison to the lower-energy protons that the
radio emission is sensitive to \citep[see][for an extended discussion of this
point]{2011arXiv1111.5544M}.

@@

Recently, \citet{2012arXiv1207.3025B} pointed out problems of the hadronic model
for the limited parameter space $\alpha\gtrsim2.6$, while adopting a much
more extended azimuthally averaged RH profile at 352~MHz
\citep{2011MNRAS.412....2B} than the one used in the present study so far
\citep{1997A&A...321...55D}. The problem of reproducing the spectral steepening
claimed in the Coma RH is somewhat mitigated when considering the negative flux
bowl due to the SZ effect that causes the measured RH spectrum to experience a
sharp cutoff towards high frequencies \citep[][which both use X-ray inferred
pseudo-pressure profiles and assume an accretion shock radius to estimate the
Compton-$y$ parameter of Coma]{2002A&A...396L..17E, 2004A&A...413...17P}. To
overcome this uncertainty, we use the observed SZ decrement of the Coma cluster
by the \citet{2012arXiv1208.3611P}. We obtain the total SZ flux by integrating
the Compton-$y$ profile up to the RH's radial extent at 352~MHz of about
$1.3$~$h_{0.7}^{-1}$~Mpc \citep[][which provide the most sensitive observation
of the Coma RH to date]{2011MNRAS.412....2B}. We show the resulting SZ decrement
in Fig.~\ref{fig:SZ_Coma} together with the Coma RH flux at different
frequencies \citep{2003A&A...397...53T}. An intrinsic power-law radio spectrum,
$(\nu/\nu_{0})^{-\alpha/2}$, exhibits a sharp cutoff at frequencies around 3-5
GHz due to the negative flux contribution of the SZ effect. Hence, we fit the
resulting RH spectrum with different spectral indices, $2.1\lesssim
\alpha\lesssim2.9$. We allow the normalization and $\nu_{0}$ to vary freely and
impose a lower floor to the reported uncertainties of 10\% to account for
possible systematics. The resulting reduced-$\chi^{2}$ values (shown in the
legend of Fig.{\ }\ref{fig:SZ_Coma}) underline the discrepancy between RH flux
measurements at different frequencies---in particular for interferometric
measurements that have to reconstruct a large fraction of the flux due to the
``missing zero spacing'' problem. At $\sim5$~GHz, the SZ decrement is comparable
to the measured RH flux. {\bf Additionally, we want to stress once more that
CR streaming can also cause spectral steepening \citep{2013arXiv1303.4746W}.}  
%This demonstrates that the magnitude of the SZ effect
%is higher than assumed by \cite{2004JKAS...37..493B}, \cite{2004A&A...424..773R}, 
%and \cite{2010MNRAS.401...47D}, while it is lower than the models of 
%\cite{2002A&A...396L..17E} and \cite{2004A&A...413...17P}.

@@

%\begin{enumerate} 
%\item 
(i) Merging clusters are not spherically symmetric as can be seen in Coma and
  Abell~2163, requiring inherently non-spherical modelling. In order to
  reproduce the more extended radio emission relative to the thermal X-ray
  emission, the non-thermal clumping factor, $C_{\rmn{non-th}}$, needs to be
  larger than its thermal analogue, $C_{\rmn{th}}$, in concentric spherical
  shells, where we defined those statistics by
  \begin{eqnarray}
    \label{eq:clumping}
    C_{\rmn{non-th}} &=&
    \expval{\rho_\rmn{gas} C}/\expval{\sqrt{\rho_\rmn{gas} C}}^2,\\
    C_{\rmn{th}} &=& 
    \expval{\rho_\rmn{gas}^2}/\expval{\rho_\rmn{gas}}^2.
  \end{eqnarray}
  This manifests itself, e.g., in the large-scale morphology of the radio
  surface brightness emission, which is more elongated than its counterpart in
  thermal X-rays, but also on scales smaller than the radio beam. In our
  phenomenological modeling, we allow for those deviations by means of the
  parameters $\gamma_{\rmn{tu}}$ and $ \alpha_{\rmn{B}}$ for the CRs and magnetic fields,
  respectively. 
  %While this approach is well suited to describe large-scale
  %anisotropies, it may be inadequate to model small-scale inhomogeneities such
  %as CR trapping in magnetic mirrors through the second adiabatic invariant and
  %needs to be carefully quantified in future work.

>>>>

  from the MultiDark $N$-body $\Lambda$CDM simulation in a companion paper.} 

@@

  a physical solution to this apparent contradiction. {\bf Assuming a redshift-independent
  fraction of radio-loud clusters, we demonstrate the unique 

@@

  LOFAR Tier~1 survey) to detect about $600\times\rmn{\emph{loud~fraction}[\%]}/10$ 
  hadronically-generated giant and mini radio halos. 
  This shows the potentially of future radio surveys in determining 

@@

\cite{2011A&A...527A..99E} tried to asses these problems of the classical hadronic 
model by analyzing CR transport processes within a cluster. While CR advection with 
turbulently driven flows results in centrally enhanced CR profiles, the propagation in 
form of CR streaming and diffusion produces a flattening of CR profiles. Hence, 
different CR transport phenomena may also account for the observed bimodality 
of the radio luminosity in the hadronic model and may also explain the spectral 
features observed in some clusters. {\bf This has been recently confirmed 
by \cite{2013arXiv1303.4746W}. In particular, they show that when considering turbulent
damping, CRs can stream at super-Alfvenic velocities, which is crucial
in order to have appreciable CR transport.} 
Note that these phenomena were not considered in earlier analytical works 
(e.g., \citealp{2004A&A...413...17P}) as well as in previous hydrodynamic 
simulations (e.g., \citealp{2001ApJ...562..233M, 
2008MNRAS.385.1211P, 2010MNRAS.409..449P}).
{\bf A satisfactory theory of CR transport in clusters does not exist yet, 
however CR streaming and diffusion may represent an intriguing solution
for the issues of the classical hadronic model.}

@@

An observational test that is able to disentangle between the hadronic and
re-acceleration models is the gamma-ray emission resulting from neutral pion
decays, a secondary product of the hadronic CR interaction with protons of the
ICM, which is not predicted by the re-acceleration model. Such observational
efforts have been undertaken in the last few years (for space-based cluster
observations in the GeV-band, see \citealt{2003ApJ...588..155R,
  2010JCAP...05..025A,2010ApJ...717L..71A,2012arXiv1207.6749H,
  2013arXiv1308.5654T,2013arXiv1308.6278H,2013arXiv1309.0197P}; for ground-based
observations in the energy band $\gtrsim100$~GeV, see
\citealt{2006ApJ...644..148P, 2008AIPC.1085..569P,
  2009arXiv0907.0727T,2009A&A...495...27A, 2009arXiv0907.3001D,
  2009arXiv0907.5000G, cangaroo_clusters, 2009ApJ...706L.275A,
  2010ApJ...710..634A, 2011arXiv1111.5544M,2012...VERITAS,2012A&A...545A.103H})
without being able to detect cluster gamma-ray emission. Current gamma-ray 
limits enable us to constrain the average CR-to-thermal pressure to be less than a few
percent, and the maximum CR acceleration efficiency at structure formation shocks 
to be $<50\%$. Alternatively, this may indeed suggest the presence of non-negligible 
CR transport processes into the outer cluster regions. 
%significant constraints on the relative CR pressure contribution. 
%The long observational campaign of the Perseus galaxy cluster
%performed by the Major Atmospheric Gamma-ray Cherenkov (MAGIC) telescopes
%constrains the cluster average CR-to-thermal pressure to be less than a few
%percent \citep{2010ApJ...710..634A,2011arXiv1111.5544M}. Comparing the
%corresponding upper limits to predictions of simulations by
%\cite{2010MNRAS.409..449P} constrains the maximum CR acceleration efficiency at
%structure formation shocks to be $<50\%$.  Alternatively, this may indeed
%suggest the presence of non-negligible CR transport processes into the outer
%cluster regions as suggested by \cite{2011A&A...527A..99E}.  Constraints at the
%same level are confirmed by the \emph{Fermi}-Large Area Telescope (LAT) data on
%the Coma cluster \citep{2011arXiv1105.3240P,2012AAS...21920701Z,
%2012...VERITAS,2012arXiv1207.6749H}. To summarize, gamma-ray observations are 
%not yet sensitive enough to severely challenge hadronic models of RHs.

@@

{\bf The main scope of this work is to improve the classical hadronic model with
the inclusion of possible CR transport phenomena, and to make predictions for
future radio surveys. The outline is as follows. In Section~\ref{sec:2.3}, we 
construct a model for the CR proton distribution in clusters that merges results 
of hydrodynamic cluster simulations and an analytical model for microscopic 
CR transport processes. In Section~\ref{sec:3}, we model observed surface 
brightness profiles of individual RHs within the hadronic scenario and explore the 
allowable parameter space for CRs and magnetic fields. Motivated by immanent 
challenges to explain the extent of giant radio halos within the hadronic model, 
we suggest a new hybrid leptonic/hadronic scenario meant to unify the apparently 
distinct classes of giant radio halos, mini halos, and steep spectrum radio sources in 
Section~\ref{sec:discussion_hadronic}. In Section~\ref{sec:4}, we apply our
extended hadronic model to a cosmologically complete mock galaxy cluster 
catalog built from the MultiDark $N$-body $\Lambda$CDM simulation
in \cite{paper1}, hereafter Paper~I. 
We compare the resulting modeled radio-to-X-ray and radio-to-SZ
scaling relations to current observations and show how they vary for different
choices of our CR and magnetic parametrization. In Section~\ref{sec:5}, we show
the radio luminosity functions, compare them to current observational
constraints and provide predictions of the hadronically-generated
RHs for the LOFAR cluster survey. Finally, in Section~\ref{sec:6}, we present 
our conclusions.} In this work, the cluster mass
$M_{\Delta}$ and radius $R_{\Delta}$ are defined with respect to a density that
is $\Delta=200$ or $\Delta=500$ times the \emph{critical} density of the
Universe. We adopt density parameters of $\Omega_{\rmn{m}}=0.3$,
$\Omega_{\Lambda}=0.7$ and today's Hubble constant of $H_0 = 100 \times
h_{70}$~km~s$^{-1}$~Mpc$^{-1}$ where $h_{70} = 0.7$.

@@

{\bf To construct such a model, we have to generalize the approach proposed by
\citet{2011A&A...527A..99E}, which uses a $\beta$-profile gas parameterization, 
in order to account for different kind of ICM gas profiles, such as our generalized 
Navarro-Frank-White (GNFW) ICM profiles derived in Paper~I from X-ray observations.} 
We also have to include the cluster mass-scaling of the
CR normalization obtained from simulations \citep{2010MNRAS.409..449P}. While
details are given in Appendix~\ref{app:B}, we show below the main steps. When
turbulent advection completely dominates the CR transport, the CR normalization
can be written as \citep{2011A&A...527A..99E}
\begin{equation}
C_{\rmn{adv}}(R)=C_{0} \left( \frac{P_{\rmn{th}}(R)}{P_{\rmn{th},0}} \right)^{\frac{\beta_{\rmn{CR}}}{\gamma}} = 
C_{0} \eta(R)^{\beta_{\rmn{CR}}},
\label{eq:Csimple_1}
\end{equation} 
where $P_{\rmn{th}}$ is the thermal pressure, $\beta_{\rmn{CR}}=(\alpha+2)/3$, $\gamma=5/3$, 
and we introduced the advective CR profile $\eta(R)=(P_{\rmn{th}}(R)/P_{\rmn{th},0})^{1/\gamma}$. Solving the continuity
equation for CRs, \citet{2011A&A...527A..99E} derive the CR density profile,
\begin{equation}
\rho_{\rmn{CR}}(R) = \rho_{\rmn{CR},0} \eta(R) \rmn{exp} \left( \frac{R}{R_{*}} \right) \, ,
\label{eg:rhoCR_1}
\end{equation} 
where $R_{*}=\gamma_{\rmn{\rmn{tu}}}R_{\rmn{c}}$ and $R_{\rmn{c}}$ is the characteristic 
radius, of order the core radius, at which the turbulence is supposed to be injected.
Now, we introduce the \emph{semi-analytical} mass-dependent normalization of the
CR profile of \cite{2010MNRAS.409..449P} such that
\begin{equation}
\eta(R) = \left( \frac{C_{\rmn{adv}}(R)}{C_0} \right)^{1/\beta_{\rmn{CR}}} \to
\left( \frac{C_{\rmn{extended}}(R)}{C_0} \right)^{1/\beta_{\rmn{CR}}} \, ,
\label{eq:eta}
\end{equation} 
which effectively redefines $C_{\rmn{adv}}(R)$ by that of our extended model, i.e.,
\begin{equation}
C_{\rmn{extended}}(R) =  \tilde{C}(R)\, \frac{\rho_{\rmn{gas}}(R)}{m_\rmn{p}} \frac{T(R)}{T_0}.
\label{eq:Cf}
\end{equation} 
Here, $\rho_{\rmn{gas}}$ is the ICM gas density and $\tilde{C}(R)$ is the normalization 
CR profile of equation~(22) of \cite{2010MNRAS.409..449P}. We additionally account for the 
temperature decline toward the cluster periphery, $T(R)$, given by the fit to the universal
temperature profile obtained from cosmological hydrodynamical simulations
\citep{2007MNRAS.378..385P,2010MNRAS.409..449P} and deep {\em Chandra} X-ray
observations \citep{2005ApJ...628..655V}. Eventually, the CR profile in our
extended model is $C(R)=C_{0}(\rho_{\rmn{CR}}(R) /
\rho_{\rmn{CR},0})^{\beta_{\rmn{CR}}}$ within $R_{\pm}$ of
equation~(\ref{eq:Rpm}), with $\rho_{\rmn{CR}}$ defined by
equation~(\ref{eg:rhoCR_1}) where $C_{\rmn{extended}}$ enters through our
redefinition of $\eta$, and $C(R) = C(R_{\pm})$ for $R > R_{+}$ and $R < R_{-}$,
respectively.

@@

In the right panel of Fig.~\ref{fig:CRFinalModel}, we compare our extended model
profile with the semi-analytical advection-only case (adopting our GNFW gas
profiles and the outer temperature decrease to the model proposed by
\citealp{2010MNRAS.409..449P}) and with the exact analytical solution as in
\citet{2011A&A...527A..99E}, but for our GNFW profiles (see Appendix~\ref{app:B}
for details).  The profiles are normalized at $0.1 R_{200}$. In the case of
dominant advective CR transport, our extended model compares nicely with the
semi-analytical model derived from cosmological cluster simulations
\citep{2010MNRAS.409..449P}.  The main differences between our extended model (and
the semi-analytical model) on the one side and the analytical solution on the
other side is the inclusion of the simulation-based ``reference'' profile
$\tilde{C}$ for the advection-only case and the universally observed temperature
drop towards the outskirts of clusters. Note that the profiles in our extended
model are generally more centrally peaked in comparison to the analytical GNFW
case, which is due to the enhanced radiative cooling in the
\citet{2010MNRAS.409..449P} simulations that did not account for AGN
feedback. Thanks to the flexible parametrization in our model, this can be
easily counteracted by changing $\gamma_{\rmn{tu}}$ and $ \alpha_{\rmn{B}}$ 
(representing the magnetic field radial decline, see next section), however, at
the expense that these parameters are now degenerate with our assumptions  on the
CR profile in the advection-dominated regime and other possible effects that we
are not considering, such as cluster asphericity.

@@

For the purpose of this section, we adopt the
measured gas and temperature profiles derived from X-ray observations of each cluster.  
Our extended model includes an overall normalization $g_{\rmn{CR}}$ of the CR
distribution function and the hadronically-induced non-thermal emission 
(Appendix~\ref{app:A}). Note that
this parameter can be interpreted as a functional that depends on the
\emph{maximum CR acceleration efficiency}, $g(\zeta_{\rmn{p,max}})$,
but \emph{only} for $\gamma_{\rmn{tu}}\gtrsim100$ \citep{2010MNRAS.409..449P}.

@@

The gamma-ray flux (Appendix~\ref{app:C}) within $R_{200}$ for the parameter
combination $\gamma_{\rmn{tu}} = 2$ and $\alpha_{\rmn{B}}=0.5$ and for energies
above 100~MeV (100~GeV) is $F_{\gamma} = 2.4 \times 10^{-9}$ ($8.7 \times
10^{-13}$) cm$^{-2}$~s$^{-1}$. 
{\bf We can compare our $F_{\gamma} (>500~\rmn{MeV}) = 6.9 \times10^{-10}$~cm$^{-2}$~s$^{-1}$
to the upper limit recently set with the \emph{Fermi} data of $F_{\gamma, UL} (>500~\rmn{MeV}) 
= 4 \times 10^{-10}$~cm$^{-2}$~s$^{-1}$ \citep{2013arXiv1308.5654T}.  
However, we note that the latter has been obtained using the \cite{2010MNRAS.409..449P}
model which is significantly more peaked than the $\gamma_{\rmn{tu}} = 2$ case
and, therefore, is too stringent for a proper comparison.}
Note also that for slightly higher values of $\gamma_{\rmn{tu}}$, i.e., a more centrally
concentrated CR distribution, the radio and gamma-ray yield would be increased
(assuming CR number conservation). However, in order to match the observed radio
synchrotron profiles, we have to decrease the CR normalization (parametrized by
$g_{\rmn{CR}}$). This causes the associated gamma-ray flux also to be reduced to
a level that is low enough to easily circumvent the gamma-ray constraints. 
{\bf E.g., for the parameter combination $\gamma_{\rmn{tu}} = 3$ and
$\alpha_{\rmn{B}}=0.4$, we obtain $F_{\gamma} = 1.3 \times 10^{-9}, 3.9 \times 10^{-10}$,
and $4.9 \times 10^{-13}$~cm$^{-2}$~s$^{-1}$ for energies above 100~MeV, 500~MeV,
and 100~GeV, respectively.}  We caution that in principle, CR streaming
should cause the CR spectrum to steepen \citep{2013arXiv1303.4746W}.  
This may then considerably weaken these constraints as a result of the convex spectral 
curvature since the gamma-ray emission probes the high-energy tail of the CR distribution 
that is suppressed in this picture in comparison to the lower-energy protons that the
radio emission is sensitive to \citep[see][for an extended discussion of this
point]{2011arXiv1111.5544M}.

@@

Recently, \citet{2012arXiv1207.3025B} pointed out problems of the hadronic model
for the limited parameter space $\alpha\gtrsim2.6$, while adopting a much
more extended azimuthally averaged RH profile at 352~MHz
\citep{2011MNRAS.412....2B} than the one used in the present study so far
\citep{1997A&A...321...55D}. The problem of reproducing the spectral steepening
claimed in the Coma RH is somewhat mitigated when considering the negative flux
bowl due to the SZ effect that causes the measured RH spectrum to experience a
sharp cutoff towards high frequencies \citep[][which both use X-ray inferred
pseudo-pressure profiles and assume an accretion shock radius to estimate the
Compton-$y$ parameter of Coma]{2002A&A...396L..17E, 2004A&A...413...17P}. To
overcome this uncertainty, we use the observed SZ decrement of the Coma cluster
by the \citet{2012arXiv1208.3611P}. We obtain the total SZ flux by integrating
the Compton-$y$ profile up to the RH's radial extent at 352~MHz of about
$1.3$~$h_{0.7}^{-1}$~Mpc \citep[][which provide the most sensitive observation
of the Coma RH to date]{2011MNRAS.412....2B}. We show the resulting SZ decrement
in Fig.~\ref{fig:SZ_Coma} together with the Coma RH flux at different
frequencies \citep{2003A&A...397...53T}. An intrinsic power-law radio spectrum,
$(\nu/\nu_{0})^{-\alpha/2}$, exhibits a sharp cutoff at frequencies around 3-5
GHz due to the negative flux contribution of the SZ effect. Hence, we fit the
resulting RH spectrum with different spectral indices, $2.1\lesssim
\alpha\lesssim2.9$. We allow the normalization and $\nu_{0}$ to vary freely and
impose a lower floor to the reported uncertainties of 10\% to account for
possible systematics. The resulting reduced-$\chi^{2}$ values (shown in the
legend of Fig.{\ }\ref{fig:SZ_Coma}) underline the discrepancy between RH flux
measurements at different frequencies---in particular for interferometric
measurements that have to reconstruct a large fraction of the flux due to the
``missing zero spacing'' problem. At $\sim5$~GHz, the SZ decrement is comparable
to the measured RH flux. {\bf Recently, \cite{2013arXiv1309.1820B}, using the
correlation between $Y_{\rmn{SZ}}$ and the low-frequency radio-halo 
brightness \citep{2011MNRAS.412....2B,2012arXiv1208.3611P}, 
found that the SZ decrement at $\sim5$~GHz is below 10\%.
This is clearly in contrast with the results from our simplified approach,
at least for $\alpha \ge 2.7$. However, as pointed out by \cite{2013arXiv1309.1820B}, 
the low-frequency observations by \cite{2011MNRAS.412....2B} imply that the Coma
radio halo is significantly larger than what thought before. This could dramatically
impact the estimations regarding the SZ decrement and more sensitive
observations are needed to better clarify this issue. On this, we want to stress once more 
that CR streaming can also cause spectral steepening \citep{2013arXiv1303.4746W}.
}

%This demonstrates that the magnitude of the SZ effect
%is higher than assumed by \cite{2004JKAS...37..493B}, \cite{2004A&A...424..773R}, 
%and \cite{2010MNRAS.401...47D}, while it is lower than the models of 
%\cite{2002A&A...396L..17E} and \cite{2004A&A...413...17P}.

@@

%\begin{enumerate} 
%\item 
(i) Merging clusters are not spherically symmetric as can be seen in Coma and
  Abell~2163, requiring inherently non-spherical modeling. In order to
  reproduce the more extended radio emission relative to the thermal X-ray
  emission, the non-thermal clumping factor, $C_{\rmn{non-th}}$, needs to be
  larger than its thermal analogue, $C_{\rmn{th}}$, in concentric spherical
  shells, where we defined those statistics by
  \begin{eqnarray}
    \label{eq:clumping}
    C_{\rmn{non-th}} &=&
    \expval{\rho_\rmn{gas} C}/\expval{\sqrt{\rho_\rmn{gas} C}}^2,\\
    C_{\rmn{th}} &=& 
    \expval{\rho_\rmn{gas}^2}/\expval{\rho_\rmn{gas}}^2.
  \end{eqnarray}
  This manifests itself, e.g., in the large-scale morphology of the radio
  surface brightness emission, which is more elongated than its counterpart in
  thermal X-rays, but also on scales smaller than the radio beam. In our
  phenomenological modeling, we allow for those deviations by means of the
  parameters $\gamma_{\rmn{tu}}$ and $ \alpha_{\rmn{B}}$ for the CRs and magnetic fields,
  respectively. 
  %While this approach is well suited to describe large-scale
  %anisotropies, it may be inadequate to model small-scale inhomogeneities such
  %as CR trapping in magnetic mirrors through the second adiabatic invariant and
  %needs to be carefully quantified in future work.
